---
title: "Spin SASmarkdown"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Use of SASmarkdown}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  \usepackage[utf8]{inputenc}
---

```{r backtick, echo=FALSE, results="hide"}
backtick <- "`"
```
`spinsas` is an alternative way of processing dynamic Markdown
from a SAS command file.  In this form, documents are written
within the comments of a SAS file.  This enables you to
run the same file within SAS (to simply execute the commands) and
through `knitr` (to produce Markdown and HTML files).

This is particularly useful if you already have a number of \*.sas
files to which you\'d like to add a little text to produce a
document.

## Special Markup
Within your SAS command file (\*.sas), you use the following symbols
to denote plain Markdown text, SAS code chunks, R code chunks, 
and text to be
dropped from the document.

Notice that SAS will treat all of these as comments.

- "** ", a double asterisk, signals the beginning of document text.  
End with a semi-colon, ";" at the end of a line.
- "*+ ", an asterisk-plus, signals the beginning of a code chunk, and
specifies the code chunk options.  End with a semi-colon.
- "*R ", an asterisk-R, signals the enclosed code is in R.  End with a semi-colon.
- "/** ", comments to "spin" begin with slash-double-asterisk.
These do not show up in your document.  End with "*/*" at
the end of a line.

## Semi-colons

- No semi-colons ***at the ends of lines*** in your document text.
- Your chunk of SAS code must end with a semi-colon (not a block comment).

## Example
  In order to run SAS code, first we specify a path for SAS and set up some options in R.
```
*+ setup, message=FALSE ;

*R 
library(SASmarkdown)
sasexe <- 'C:/Program Files/SASHome/SASFoundation/9.4/sas.exe'
sasopts <- '-nosplash -ls 75'
;
```
Text is then included as a special comment.
```
** The report begins here.;
```

Finally, the executable SAS code is given a line of chunk instructions.

```
*+  example1, engine='sas', engine.path=saspath, engine.opts=sasopts, 
comment=NA;

proc means data=sashelp.class /*(keep = age)*/;
run;

/* lines here are
ignored by SAS.  If at the end of
your SAS code chunk they must be followed by a semi-colon.
*/
;
```

You can use the ***usual*** Markdown within the text sections.

The entire document might be:
```
*+ setup, message=FALSE ;

*R 
library(SASmarkdown)
sasexe <- 'C:/Program Files/SASHome/SASFoundation/9.4/sas.exe'
sasopts <- '-nosplash -ls 75'
;

** The report begins here.;

*+  example1, engine='sas', engine.path=sasexe, engine.opts=sasopts, 
comment=NA;

proc means data=sashelp.class /*(keep = age)*/;
run;

```

## Processing
Set up an example file to use:
```{r ex, comment=NA, results="hide"}
indoc <- "
*+ setup, message=FALSE ;

*R 
library(SASmarkdown)
sasexe <- 'C:/Program Files/SASHome/SASFoundation/9.4/sas.exe'
sasopts <- '-nosplash -ls 75'
;

** The report begins here.;

*+  example1, engine='sas', engine.path=sasexe, engine.opts=sasopts, 
comment=NA;

proc means data=sashelp.class /*(keep = age)*/;
run;
"
writeLines(indoc, "indoc.sas")
```

To process this document then, simply use

```{r spin, comment=NA, results="hide"}
library(SASmarkdown)
spinsas("indoc.sas")
```

Which gives a document, \"indoc.html\" that looks like

<hr>
```{r output, child="indoc.html"}
```

```{r cleanup, echo=FALSE, results="hide"}
unlink("indoc.sas")
unlink("indoc.md")
unlink("indoc.html")
```