---
title: "SAS Using R Markdown (Windows)"
author: "Doug Hemken"
date: "May 2017"
output:
  html_document:
    highlight: null
    self_contained: no
    theme: null
    toc: yes
  pdf_document:
    toc: yes
---


```{r}
  require(SASmarkdown)
  backtick <- "`"
  engine_output <- sasengine_output()
  sashtmlengine()
  saslogengine()
  sas_collectcode()
  source("./R/SasMDsetup.r")
```

## Introduction
This is an introduction to using R Markdown to produce
documents based on SAS.  

Using SAS with R Markdown gives you a method of writing simple documents that
include the results of executing SAS commands, with both the document text 
and the source SAS code in one file.

This is especially good for writing simple documents that explain statistical
coding tasks.  One way to think of this is as an extended and more 
readable version of writing
comments in your source code.

This process uses 
Rstudio (or just R) to create documents that depend upon SAS code.
The source for **this** document is [SASmarkdown.rmd](SASmarkdown.rmd)

## Background
*Markdown* is a language for formatting fairly simple documents using 
just a few text symbols.  It is designed to be easy to read and write. If 
you are not already familiar with Markdown see John Gruber\'s
[Markdown article](https://daringfireball.net/projects/markdown/). 

*R Markdown* extends Markdown by allowing you to include blocks of code in one
of several programming languages.  The code is evaluated, and both the
code and it\'s results are included in a Markdown document.  To read
more about the details of R Markdown see RStudio\'s 
[R Markdown webpages](http://rmarkdown.rstudio.com/lesson-1.html)

*RStudio* uses an R package called `knitr` (this could also be called
directly from R), which includes the ability to evaluate SAS code.  We
will refine this in subsequent articles.

The documentation for `knitr` can be found in R\'s Help, from 
[this web page](https://yihui.name/knitr/), or in Yihui Xie\'s book,
*Dynamic Documents with R and Knitr*.

RStudio makes
it easy to process your R Markdown document, execute the source code, and produce
a final HTML, pdf, or Word
document.  After you have written your R Markdown document, RStudio processes
it with a simple click of a button!  The focus in these articles will be
on producing HTML (web) pages.

Using SAS with R Markdown is not quite as graceful as using R 
with R Markdown, but we have automated some techniques for putting together
simple documents.

Note that this is not a friendly environment for extensively
debugging problems in your SAS code.  If your code
is at all complicated, you should work out the details
in SAS first, then bring it into RStudio to develop
your documentation!

## Including Your SAS Code
SAS code is included in your R Markdown document in a block called
a \"code chunk\".  When you click on Rstudio\'s Knit button, your 
initial document, what you actually write out,
is processed by the R function `knitr`, evaluating your code, collecting
the output, and producing a Markdown document. This Markdown document
is then processed by a program called *Pandoc* to produce your final
HTML document.

It helps to understand how knitr evaluates your SAS code.  Knitr writes
out your code into a file, and then submits that file to SAS to be
run in batch mode.  This produces SAS log and SAS \"listing\" output.
The listing output is then read back into knitr and used to produce
an block of output in the Markdown document.  (These files are all
temporary.)

### Basic Limitations
There are two basic limitations that you will have to work with:

* Each code chunk runs separately as a SAS batch job, they 
are not processed as a
single running session like R code.  Interspersing text with executed
code is not as simple as when running R code.  

    For example, suppose you
have a DATA step early in your document, and then want to use that
data in a later code chunk.  In the later code chunk,
SAS will have begun a new session (a new batch job) and \"forgotten\" the first
DATA step.  SAS will not find the data, so SAS will send an error message to
R Markdown.  The message R Markdown passes back to you will only tell you
that SAS did not exit properly.

    (Unlike calling R from 
SAS\'s `PROC IML` (see the 
[*Calling R from SAS* KB](../SASWindows/RfromSAS.html)),
when calling SAS from R you cannot just hand off your data).

    Some of this difficulty
can be overcome with the strategic use of a SAS autoexec file (which
we automate in [another article](SASmdlinkcode.html)) or
by echoing only part of the code through the `echo` code chunk option.

* If you want to use SAS\'s html output instead of the listing
output, or if you want to include SAS graphics in your final document,
a different version of knitr\'s SAS engine makes this fairly simple
(again, a topic for [another article](SAShtmlengine.html)).

### HTML or PDF as a Final Target?
Longtime SAS users will find that SAS listing output looks as
expected in both html and pdf documents.  (Newer SAS users may not be
familiar with listing output, since it is no longer the default when
using SAS interactively.  R Markdown runs SAS in batch mode,
so listing output is produced by default.)

SAS html output can be included in a final html document, but renders
poorly as a pdf document when run through R Markdown.  To do much more 
than simply display listing output 
in pdf documents, you will want to investigate SAS\'s 
[StatRep](http://support.sas.com/resources/papers/proceedings12/324-2012.pdf) 
or Lenth\'s [SASWeave](http://homepage.cs.uiowa.edu/~rlenth/SASweave/).

### File locations
Because I am running this on my office (SSCC) desktop and on Winstat,
everything goes on the Z: drive.  As long as I specify
the root of Z: in the `engine.opts`, both
R Markdown and SAS default to the folder where the
markdown file is saved.  

(I don\'t have this fully worked out for non-SSCC scenarios.
If you are working
in another environment, you will have to be more careful about
where SAS writes its output files, and where R looks for them.)

(Also keep in mind that for the Windows OS, SAS will use backslashes
in file paths, while R will use forward slashes.  Note in the **code
chunk** options below we are specifying paths to R, so use forward slashes.)

### Code Setup
The code chunk setup to run SAS is not too difficult.  But as long as 
SAS is
not in your operating system\'s `PATH` variable, you need to specify its
full path with the `engine.path` option (all on one line), e.g.

    `r backtick```{r, engine="sas", engine.path="C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"}
    
    -- SAS code here --
    
    `r backtick```

You will also want to add an `engine.opts` option, so that SAS will
save the listing output (and the log) to a location where R will handle
it gracefully, including it in the final document automatically and
deleting the SAS output file when finished, e.g.

    `r backtick```{r, engine='sas', engine.path="C:/Program Files/SASHome/SASFoundation/9.4/sas.exe", engine.opts="-nosplash -log 'z:' -print 'Z:'", comment=""}
    
    -- SAS code here --
    
    `r backtick```
    
If you have several code chunks you can automate this with a preliminary
code chunk in R as

    `r backtick```{r}
    saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
    sasopts <- "-nosplash -log 'z:' -print 'Z:' -ls 75"
    `r backtick```

    `r backtick```{r, engine='sas', engine.path=saspath, engine.opts=sasopts, comment=""}
    
    -- SAS code here --
    
    `r backtick```
    
Finally, if all of your executable code is sent to SAS  you can 
just set the chunk
options as defaults:

    `r backtick```{r, message=FALSE}
    require(knitr)
    saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
    sasopts <- "-nosplash -log 'z:' -print 'Z:' -ls 75"
    opts_chunk$set(engine="sas", engine.path=saspath, 
        engine.opts=sasopts, comment="")
    `r backtick```

    `r backtick```{r}
    
    -- SAS code here --
    
    `r backtick```

## Running SAS
Here then, is an example as it might appear in your final document.
```{r, engine='sas', engine.path=saspath, engine.opts=sasopts, comment=""}
proc means data=sashelp.class;
run;

proc freq data=sashelp.class;
    tables sex;
run;
```

## HTML Output and Graphics
There are two advantages to using SAS\'s html output, rather than
the listing output.  First, results tables become html tables instead
of strategically spaced text, and will dynamically size
themselves to the end user\'s window.  Second, you can include
SAS graphics in your final document with a minimum of fuss.

To use SAS\'s html output, use `sashtml` engine, as described in a later
article.

    `r backtick```{r, engine='sashtml', engine.path=saspath, engine.opts=sasopts, results='asis'}
    -- SAS code here --
    `r backtick```

As an example,
```{r, engine='sashtml', engine.path=saspath, engine.opts=sasopts, results='asis'}
proc corr data=sashelp.class plots=matrix;
run;

