---
title: "Basic Use of SASmarkdown"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Basic Use of SASmarkdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init, include=FALSE}
try(detach("package:SASmarkdown"), silent=TRUE)
```

This discussion assumes you have a basic understanding
of Markdown for document formatting, Rmarkdown to include executable
code in a document, and SAS to write the code.

SASmarkdown defines six knitr engines, to return different
combinations of SAS code, SAS listing output, SAS
HTML output, and SAS log output for use in your document.

SASmarkdown also defines several knitr chunk (code block) options
for working with SAS output.

## Set up your SAS engine configuration.
The details depend on your operating system, the version of SAS, and
whether or not you have SAS installed in a default location.

    ```{r setup, message=FALSE} `r ''`
    library(SASmarkdown)
    ```

```{r setup, include=FALSE}
library(SASmarkdown)
```

This defines six knitr engines

- *sas* - SAS code and listing output
- *saslog* - SAS log and listing output
- *sashtml* - code and HTML output
- *sashtmllog* - log and HTML output
- *sashtml5* - code and HTML5 output
- *sashtml5log* - log and HTML5 output

These are ordinarily used in pairs, to give you access to
the code, log, and output format you prefer.

**If SAS is not found**, you will have to define the engine path
and any engine options in your document.  For example

    ```{r enginepath} `r ''`
    saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
    sasopts <- "-nosplash -ls 75" # '-nosplash' fails in some unix terminals
    knitr::opts_chunk$set(engine.path=list(sas=saspath, saslog=saspath),
                      engine.opts=list(sas=sasopts, saslog=sasopts), 
                      comment=NA)
    ```
## Set up SAS code chunks.
A simple code chunk in your document might look like:

    ```{sas example1} `r ''`
    proc means data=sashelp.class (keep = age);
    run;
    ```

And in your document this would produce:
```{sas example1}
proc means data=sashelp.class (keep = age);
run;
```

## HTML output.
HTML gives you a method of embedding SAS tables and graphs
directly in your document.

Switch the engine specification to **sashtml**.  This example 
suppresses the code echo with the `echo=FALSE` chunk option.

    ```{sashtml example2, echo=FALSE} `r ''`
    proc means data=sashelp.class (keep = age);
    run;
    ```

Which produces:
```{sashtml example2, echo=FALSE}
proc means data=sashelp.class (keep = age);
run;
```

## HTML with graphics.
If you use the `sashtml` engine, nothing special is required to
include SAS ODS graphics.

    ```{sashtml example3} `r ''`
    proc corr data=sashelp.class nosimple plots=matrix;
    run;
    ```

Producing:
```{sashtml example3}
proc corr data=sashelp.class nosimple plots=matrix;
run;
```

## Show SAS log files.
We can repeat the first example, showing the SAS log instead of
the SAS code by using the **saslog** engine.

    ```{saslog example4} `r ''`
    ```

Producing:
```{saslog example4}
proc means data=sashelp.class (keep = age);
run;
```

## Repeat with both log and html output.
Finally, you can have both the SAS log and the HTML output with
the **sashtmllog** engine.

    ```{sashtmllog example1} `r ''`
    ```

Producing:
```{sashtmllog example1}
```
