---
title: "Linking SAS Code Chunks"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linking SAS Code Chunks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r init, include=FALSE}
try(detach("package:SASmarkdown"), silent=TRUE)
```

Linking SAS code chunks makes use of SAS\'s autoexec.sas
file.  When you attach the SASmarkdown library, a chunk
option is set up to place chunks in an autoexc file.

## Configure the SAS engine
```{r setup, message=FALSE}
library(SASmarkdown)
```

## Set up to carry over
To have the result of one code chunk available to use in a later
code chunk, set the chunk option `collectcode=TRUE`.

In this example, a data set is copied to the WORK library
and only one variable is kept.

In Rmarkdown this would look like:

    ```{sas datastep, collectcode=TRUE} `r ''`
    data class;
        set sashelp.class;
        keep age;
        run;
    ```

And in your final document it would appear as:
```{sas datastep, collectcode=TRUE}
data class;
    set sashelp.class;
    keep age;
    run;
```

## Using the previous code chunk
Without **collectcode** to link the code chunks, a later
chunk that referenced the data in the WORK library would
produce an error, but this now works.  (No special option
is needed for this later step.)
```{sas procmeans, error=TRUE}
proc means data=class;
run;
```
