---
title: "Linking Code Chunks Using 'collectcode'"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linking Code Chunks Using collectcode}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---

```{r backtick, include=FALSE}
backtick <- "`"
```
Linking SAS code chunks makes use of SAS\'s autoexec.sas
file.  When you attach the SASmarkdown library, a chunk
option is set up to place chunks in an autoexc file.

## Configure the SAS engine
```{r setup, message=FALSE}
library(SASmarkdown)
saspath <- find_sas()
sasopts <- "-nosplash -ls 75"

knitr::opts_chunk$set(engine.path=list(sas=saspath, saslog=saspath),
                      engine.opts=list(sas=sasopts, saslog=sasopts),
                      comment=NA)
```

## Then set up SAS code chunks.
To have the effects of one code chunk available to use in a later
code chunk, set the chunk option \"collectcode=TRUE\".

In this example, a data set is copied to the WORK library
and only one variable is kept.

In Rmarkdown this would look like:

    ```{sas datastep, collectcode=TRUE}`r `
    data class;
        set sashelp.class;
        keep age;
        run;
    `r backtick```

And in your final document it would appear as:
```{sas datastep, collectcode=TRUE}
data class;
    set sashelp.class;
    keep age;
    run;
```
## A later step - using the WORK data.
Without **collectcode** to link the code chunks, a later
chunk that referenced the data in the WORK library would
produce an error, but this now works.  (No special option
is needed for this later step.)
```{sas procmeans, error=TRUE}
proc means data=class;
run;
```
