---
title: "Showing SAS errors"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Showing SAS errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r detach, include=FALSE}
try(detach("package:SASmarkdown"), silent=TRUE)
```

SAS does not always exit with an error when it encounters problems, but
when it does, you should see the log in your document instead of the code.

## First, set up your SAS engine configuration.
```{r setup, message=FALSE}
library(SASmarkdown)
```

## A Semantic Error
Semantic errors (typos etc.) cause SAS to return an error
status when SAS finishes.  This will automatically cause 
SASmarkdown to switch from the `sas` engine to the `saslog`
engine.

In this example, trying to use WORK data without any data there,
SAS produces an ERROR, and no output is produced.)

    ```{sas procmeans} `r ''`
    proc means data=class;
    run;
    ```

Which produces:
```{sas procmeans}
proc means data=class;
run;
```
## An Execution Error.
Division by zero produces error flags within the DATA step, but does
**not** cause SAS to return an error code when it finishes.  The `sas`
engine gives us the usual output.

To see the DATA step ERROR, switch to the `saslog` engine.

    ```{saslog divzero} `r ''`
    
    *--- sas code ---;
    
    ```

Producing both the log and the PROC PRINT output.
```{sas divzero}
data class;
    set sashelp.class(obs=5);
    age0 = age/0;
    keep age age0;
    run;

proc print data=class;
run;
```


## Replicate in HTML
For a semantic error, which produces no output tables or graphs,
the **sas** and **sashtml** engines give us the
same result.
```{sashtml procmeans, error=TRUE}
```

For an execution error, using 
the **sashtmllog** engine, we see both the DATA step errors and the
HTML output.
```{sashtmllog procmeans, error=TRUE}
```
