---
title: "Cleaning Up SAS Log Output"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning Up SAS Log Output}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
library(SASmarkdown)
```


If you are using the **saslog** or **sashtmllog** engines, or if you
are allowing SAS semantic errors to appear in your document, you may
not want to show the entire log file.  To gain some control over
what shows up in your document, you can use chunk options.


The usual SAS set up is needed, as in other vignettes, and is not
shown here.

### Hook Specification
These options are set up for you when you load the SASmarkdown library.


### Hook Use
First, we set up some code using the `sas` engine.  We will
repeatedly use the same code by repeating the chunk label.
```{sas procmeans2}
proc means data=sashelp.class(keep=height);
run;
```

If we switch to engine **saslog**

    ```{saslog procmeans2} `r ''`
    ```

we don't see ALL of the SAS log file, but we may be showing
more information than we really want to present to the reader.
```{saslog procmeans2}
```

We can filter the log to suppress processing times by 
using chunk option `SASproctime=FALSE`.

    ```{saslog procmeans2, SASproctime=FALSE, results='hide'} `r ''`
    ```

```{saslog procmeans2, SASproctime=FALSE, results='hide'}
```

Other filtering options include no SAS command echo with
`SASecho=FALSE` and no SAS NOTES with `SASnotes=FALSE`.

```{saslog procmeans2, SASecho=FALSE, results="hide"}
```

```{saslog procmeans2, SASnotes=FALSE, results="hide"}
```

### Interactions (or Lack Thereof)
Chunks in other languages are unaffected (for example, R).
```{r sourcetest1}
runif(5)
```

And the **sas** and **sashtml** engines are unaffected, as seen
in our intial code chunk and here:
```{sashtml procmeans2}
```
