---
title: "Cleaning Up SAS Log Output"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning Up SAS Log Output}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  \usepackage[utf8]{inputenc}
---
```{r backtick, echo=FALSE, results="hide"}
backtick <- "`"
```

```{r setup, echo=FALSE, message=FALSE}
require(SASmarkdown)
if (file.exists("C:/Program Files/SASHome/SASFoundation/9.4/sas.exe")) {
  saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
} else {
  saspath <- "sas"
}
sasopts <- "-nosplash -ls 75"
knitr::opts_chunk$set(engine.path=saspath, engine.opts=sasopts, comment="")
```

If you are using the **saslog** or **sashtmllog** engines, or if you
are allowing SAS semantic errors to appear in your document, you may
not want to show the entire log file.  To gain some control over
what shows up in your document, you can use \"saslog_hookset\".

The usual SAS set up is needed, as in other vignettes, and is not
shown here.

## A Source Hook for the \"saslog\" Engine
With the **saslog** or **sashtmllog** engines we use the 
\"saslog_hookset\" function, specifying  
a *source* hook.  This creates a number of chunk options we can
use to filter the SAS log.

### Hook Specification
(This is a \"source\" hook, because we are showing the SAS log
where we usually show the SAS code, the \"source\".)
```{r sourcehook}
require(SASmarkdown) # if not invoked previously
saslog_hookset("source")
```

### Hook Use
First, we setup some code using \"engine=\'sas\'\".  We will
repeatedly use the same code by repeating the chunk label.
```{r procmeans2, engine='sas'}
proc means data=sashelp.class(keep=height);
run;
```

If we switch to engine **saslog**

    `r backtick```{r procmeans2, engine='saslog'}
    `r backtick```

we don't get ALL of the SAS log file, but we may be showing
more information than we really want to present to the reader.
```{r procmeans2, engine='saslog'}
```

We can filter the log to suppress processing times by 
using chunk option \"SASproctime=FALSE\".

    `r backtick```{r procmeans2, engine='saslog', SASproctime=FALSE, results='hide'}
    `r backtick```

```{r procmeans2, engine='saslog', SASproctime=FALSE, results='hide'}
```

Other filtering options include no SAS command echo with
\"SASecho=FALSE\" and no SAS NOTES with \"SASnotes=FALSE\".

```{r procmeans2, engine='saslog', SASecho=FALSE, results="hide"}
```

```{r procmeans2, engine='saslog', SASnotes=FALSE, results="hide"}
```

### Interactions (or Lack Thereof)
Chunks in other languages are unaffected (for example, R).
```{r sourcetest1}
runif(5)
```

And the **sas** and **sashtml** engines are unaffected, as seen
in our intial code chunk and here:
```{r procmeans2, engine='sashtml'}
```

## An Output Hook for PROC PRINTTO
These same filters can be used as \"output\" hooks.  

Suppose you
wanted to process SAS log files separately.  You could use
SAS\'s PROC PRINTTO to save the log files, and read them
into your document using R commands.

(For this vignette we reset the \"source\" hook, from above.  In
general you will NOT need to reset, you will use one or the other.)
```{r outputhook, echo=TRUE}
knitr::knit_hooks$set(source=hook_orig) # just for this example

saslog_hookset("output")
```

In this example I am using the PROC PRINTTO method of capturing the log.

Note that **all** output is run through the output hook, but our alterations
are specific to the SAS log  None of the text we are filtering appears
in ordinary SAS or R output, so SAS output is untouched.

```{r procmeans, engine='sas'}
proc printto log="saslog.log" new;
proc means data=sashelp.class(keep=weight);
run;
```

The log is then processed when it is shown by a separate code chunk (using
R commands, not SAS).

    `r backtick```{r readlog}
    cat(readLines("saslog.log"), sep="\n")
    `r backtick```

```{r readlog}
cat(readLines("saslog.log"), sep="\n")
```

We can now use the same chunk options as before.  
No processing times, SASproctime=FALSE.
```{r readlog, SASproctime=FALSE, echo=FALSE}
```

No SAS command echo, SASecho=FALSE.
```{r readlog, SASecho=FALSE, echo=FALSE}
```

No SAS NOTES, SASnotes=FALSE.
```{r readlog, SASnotes=FALSE, echo=FALSE}
```

Clean up is necessary in a final R chunk - this is NOT automated.
```{r cleanuplog}
# Do not forget to remove the log file when you are done!
unlink("saslog.log")
```

