---
title: "Saving Intermediate SAS files"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Cleaning Up SAS Log Output}
  %\VignetteEngine{knitr::rmarkdown_notangle}
  %\VignetteEncoding{UTF-8}
---
```{r backtick, echo=FALSE, results="hide"}
backtick <- "`"
```

```{r setup, echo=FALSE, message=FALSE}
require(SASmarkdown)
if (file.exists("C:/Program Files/SASHome/SASFoundation/9.4/sas.exe")) {
  saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
} else {
  saspath <- "sas"
}
sasopts <- "-nosplash -ls 75"
knitr::opts_chunk$set(engine.path=saspath, engine.opts=sasopts, comment="")
```

You may want to save the intermediate SAS files.  This works with all
of the SAS engines.

The file names are taken from the code chunk label.

First, we run some code using \"engine=\'sas\'\" without saving
the intermediate SAS files.  Not shown here, this code chunk
has label \"procmeans0\".
```{r procmeans0, engine='sas'}
proc means data=sashelp.class(keep=height);
run;
```

If we invoke the \"saveSAS=TRUE\" option to save the intermediate SAS files:


    `r backtick```{r procmeans1, engine='sas', saveSAS=TRUE, results="hide"}
    proc means data=sashelp.class(keep=height);
    run;
    `r backtick```

we will save three intermediate files:  procmeans1.sas,
procmeans1.log, and procmeans1.lst.

```{r procmeans1, engine='sas', saveSAS=TRUE, results="hide"}
proc means data=sashelp.class(keep=height);
run;
```

If we stopped here you could see these in a file explorer.
Here we\'ll show them with some R code.
```{r showfiles}
list.files(pattern="proc")
```

Then finally, for this example, we will delete them.
```{r cleanup}
unlink("procmeans?.*")
```