---
title: "Linking Code Chunks Using 'sas_collectcode'"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Linking Code Chunks Using sas_collectcode}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r backtick, echo=FALSE, results="hide"}
backtick <- "`"
```
Linking SAS code chunks makes use of SAS\'s autoexec.sas
file.  Within R this requires setting up a chunk option
by invoking the **sas_collectcode** function.

## Configure SAS and call \"sas_collectcode\".
```{r setup, message=FALSE}
require(SASmarkdown)
if (file.exists("C:/Program Files/SASHome/SASFoundation/9.4/sas.exe")) {
  saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
} else {
  saspath <- "sas"
}
sasopts <- "-nosplash -ls 75"

sas_collectcode()
```
## Then set up SAS code chunks.
To have the effects of one code chunk available to use in a later
code chunk, set the chunk option \"collectcode=TRUE\".

In this example, a data set is copied to the WORK library
and only one variable is kept.

In Rmarkdown this would look like:

    `r backtick```{r datastep, engine="sas", collectcode=TRUE, engine.path=saspath, engine.opts=sasopts, comment=""}
    data class;
        set sashelp.class;
        keep age;
        run;
    `r backtick```

And in your final document it would appear as:
```{r datastep, engine="sas", collectcode=TRUE, engine.path=saspath, engine.opts=sasopts, comment=""}
data class;
    set sashelp.class;
    keep age;
    run;
```
## A later step - using the WORK data.
Without **collectcode** to link the code chunks, a later
chunk that referenced the data in the WORK library would
produce an error, but this now works.  (No special option
is needed for this later step.)
```{r procmeans, engine="sas", engine.path=saspath, engine.opts=sasopts, comment=""}
proc means data=class;
run;
```
