---
title: "Showing of SAS errors"
author: "Doug Hemken"
date: "`r format(Sys.Date(), '%d %b %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Showing of SAS errors}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---
# To see SAS errors, use chunk option \"error=TRUE\".
SAS does not always exit with an error when it encounters problems, but
when it does, you should see the log in your document instead of the code.

## In a first code chunk, set up typical SAS engine configuration.
```{r setup, message=FALSE}
require(SASmarkdown)
sas_collectcode()

if (file.exists("C:/Program Files/SASHome/SASFoundation/9.4/sas.exe")) {
  saspath <- "C:/Program Files/SASHome/SASFoundation/9.4/sas.exe"
} else if (file.exists("D:/SASHome/SASFoundation/9.4/sas.exe")) {
  saspath <- "D:/SASHome/SASFoundation/9.4/sas.exe"
} else {
  saspath <- "sas"
}

sasopts <- "-nosplash -ls 75"

```
## A Semantic Error.
Try to use WORK data, without setting up data there.
SAS produces an ERROR.  (No output is produced.)
```{r procmeans, engine="sas", engine.path=saspath, engine.opts=sasopts, comment="", error=TRUE}
proc means data=class;
run;
```
## An Execution Error.
Division by zero.
```{r divzero, engine="sas", engine.path=saspath, engine.opts=sasopts, comment="", error=TRUE}
data class;
    set sashelp.class(obs=5);
    age0 = age/0;
    keep age age0;
    run;

proc print data=class;
run;
```

To see the ERROR, switch to \"saslog\" engine.
```{r divzero, engine="saslog", engine.path=saspath, engine.opts=sasopts, comment="", error=TRUE}
```

## Replicate in HTML.
The semantic error.
```{r procmeans, engine="sashtml", engine.path=saspath, engine.opts=sasopts, comment="", error=TRUE}
```

And the execution error.
```{r divzero, engine="sashtmllog", engine.path=saspath, engine.opts=sasopts, comment="", error=TRUE}
```
